---
title: "Project 1"
author: 
- "Júlíus Pétur Guðjohnsen"
- "Sigurður Jónsson"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output: html_document
---
### Assignment 1
#### Question 1


<!-- Hérna er cóðinn fyrir fyrsta lið án evaluation -->
The following code sample shows how to import the data needed for the assignement hosted on a https server "https://raw.githubusercontent.com/juliuspeturg/Assignment1/master/fuel.csv".
To better understand the data a summary of the imported data is displayed bellow aswell as a lineplot of the fuel price index
```{r, message=FALSE,echo=TRUE,eval=TRUE}
#remove all variables af data from global enviroment
rm(list=ls())
suppressPackageStartupMessages(library(RCurl))
# load csv file from URL
x = getURL("https://raw.githubusercontent.com/juliuspeturg/Assignment1/master/fuel.csv")
fuel = read.csv(text=x,stringsAsFactors = FALSE)
summary(fuel)
attach(fuel)
par(mfrow=c(1,1))
plot(fpi~rtime,type = "l",main="Fuel price index in the US")
grid()
```

#### Question 2
For statinary time series the ACF will drop to zero very quickly however for nonstationary time series the ACF will taper off slowly. As we can see by looking at the ACF of the fuel price index the ACF tapers off slowly, from that we can conclude that the timeseries is nonstationary and therefor the mean of the series is not going to informe us a lot on the timeseries, see figure below for ACF of timeseries  
```{r eval=TRUE, echo=TRUE,message=FALSE}
#Generate timeseries object for fuelindex
fuel.ts = ts(data = fuel$fpi, start = c(1979,1),frequency = 12)
#evaluate the mean of the timeseries
cat(c("mean_fpi:",mean(fpi)))
acf(fuel.ts,main="ACF for fuel price index")
grid()
```

#### Question 3

```{r eval=TRUE, echo=TRUE,message=FALSE}
# Linear Model
mod <- lm(fpi~rtime)
summary(mod)
```

#### Question 4

```{r eval=TRUE, echo=TRUE,message=FALSE}
# Trainingset
myts <- ts(fpi, start = c(1979,1), end =c(2003,12),frequency = 12)
# Testset
myts2 <- ts(fpi, start = c(2004,1), end =c(2004,12),frequency = 12)
require(forecast)
monthplot(myts)
seasonplot(myts)
par(mfrow=c(1,1))
fit1 <- stl(myts,s.window = "period")
#3 eða 4
plot(fit1)
```

#### Question 5

```{r eval=TRUE, echo=TRUE,message=FALSE}
fuel.ts <- ts(fuel$fpi, frequency=12, start=c(1979,1))

fuel_train <- window(fuel.ts, start=c(1979,1), end=c(2003,12))
fuel_test <- window(fuel.ts, start=c(2004,1), end=c(2004,12))


```

#### Question 6

```{r eval=TRUE, echo=TRUE,message=FALSE}
fit2 <- ses(fuel_train,h = 12, alpha = 0.1,exponential=TRUE)
f.fit2 <- forecast(fit2,12)
plot(f.fit2,ylab="fpi",xlab="time")
lines(fuel_train)
lines(fuel_test,col=2)
grid()
```

#### Question 7

```{r eval=TRUE, echo=TRUE,message=FALSE}
fit3 <- ses(fuel_train,h = 12, exponential=FALSE)
f.fit3 <- forecast(fit3,12)
plot(f.fit3,ylab="fpi",xlab="time")
lines(fuel_train)
lines(fuel_test,col=2)
grid()
```

#### Question 8

```{r eval=TRUE, echo=TRUE,message=FALSE}
fit4 <- HoltWinters(myts, beta = FALSE,gamma = FALSE)
f.fit4 <- forecast(fit4,12)
plot(f.fit4,ylab="fpi",xlab="time")
lines(fuel_train)
lines(fuel_test,col=2)
grid()
```

#### Question 9

```{r eval=TRUE, echo=TRUE,message=FALSE}
fit5 <- rwf(myts, h=12, drift=FALSE, fan=FALSE, lambda=NULL, biasadj=FALSE)
f.fit5 <- forecast(fit5,12)
plot(f.fit5,ylab="fpi",xlab="time")
lines(fuel_train)
lines(fuel_test,col=2)
grid()
```

### Assignment 2
#### Question 1

```{r eval=TRUE, echo=TRUE,message=FALSE}
library(RCurl)
x = getURL("https://raw.githubusercontent.com/juliuspeturg/Assignment1/master/simdata.csv")
simdata = read.csv(text=x,stringsAsFactors = FALSE,header=TRUE, sep="")

# A2-Q1: Plot the data

attach(simdata)

# á sama grafi
plot(y1, type="l",col='red', main="Simdata time series",
     xlab="Time", ylab="Data")
lines(y2,col='blue')
lines(y3,col='green')
legend('topright', legend=c("y1", "y2","y3"),
       col=c('red','blue','green'), lty=1, bty='y', cex=.85)
```

#### Question 2

```{r eval=TRUE, echo=TRUE,message=FALSE}
par(mfrow=c(3,2)) 
# par(mfrow=c(1,1)) 
acf(y1)
pacf(y1)
acf(y2)
pacf(y2)
acf(y3)
pacf(y3)
```

#### Question 3

```{r eval=TRUE, echo=TRUE,message=FALSE}
# A2-Q3: Identify the models using arima  <<< ATH BETUR

arima(y1)
arima(y2)
arima(y3)

auto.arima(y1) # líka hægt að nota auto.arima
```

#### Question 4

```{r eval=TRUE, echo=TRUE,message=FALSE}
ts.sim <- arima.sim(list(order = c(1,0,0), ar = 0.9), n = 250)
```

#### Question 5

```{r eval=TRUE, echo=TRUE,message=FALSE}
layout(matrix(c(1,1,2,3), 2, 2, byrow = TRUE))  
plot(ts.sim, main="Simulated time series ts.sim")
acf(ts.sim, main="ACF for ts.sim")
pacf(ts.sim, main="PACF for ts.sim")
```

#### Question 6

```{r eval=TRUE, echo=TRUE,message=FALSE}
acf.plot <- function(y){
  layout(matrix(c(1,1,2,3), 2, 2, byrow = TRUE)) 
  plot(y, main=c("Simulated time series",deparse(substitute(y))))
  acf(y, main=c("ACF for",deparse(substitute(y))))
  pacf(y,main=c("PACF for",deparse(substitute(y))))
}
```

#### Question 7

```{r eval=TRUE, echo=TRUE,message=FALSE}
ts.sim1 <- arima.sim(n=250, list(ar=c(0.5)))        
ts.sim2 <- arima.sim(n=250, list(ar=c(-0.5)))
ts.sim3 <- arima.sim(n=250, list(ar=c(0.5,0.2)))
ts.sim4 <- arima.sim(n=250, list(ar=c(0.7)))
ts.sim5 <- arima.sim(n=250, list(ar=c(0.5,-0.2)))
ts.sim6 <- arima.sim(n=250, list(ma=c(0.5,0.5)))
ts.sim7 <- arima.sim(n=250, list(ma=c(-0.9)))
ts.sim8 <- arima.sim(n=250, list(ar=c(0.5), ma=c(-0.8)))

# Plotting the simulated series with the acf function
acf.plot(ts.sim1)
acf.plot(ts.sim2)
acf.plot(ts.sim3)
acf.plot(ts.sim4)
acf.plot(ts.sim5)
acf.plot(ts.sim6)
acf.plot(ts.sim7)
acf.plot(ts.sim8)
```

#### Question 8

```{r eval=TRUE, echo=TRUE,message=FALSE}
ts.sim9 <- arima.sim(list(order = c(1,1,1), ar = 0.7, ma=0.3), n = 250)

ts.sim9_diff1 <- diff(ts.sim9, differences=1) #first difference
plot.ts(ts.sim9_diff1, col=3)
lines(ts.sim9, col=2)
```
