---
title: "Project 2"
author: 
- "Júlíus Pétur Guðjohnsen"
- "Sigurður Jónsson"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output: 
  html_document: 
    fig_height: 4
    fig_width: 7
    fig_caption: yes
    self_contained: yes
    smart: no
---

```{r auto_num_functions, echo=FALSE}
# A function for generating captions and cross-references

fig <- local({
    i <- 0
    list(
        cap=function(refName, text, center=FALSE, col="black", inline=FALSE) {
            i <<- i + 1
            ref[[refName]] <<- i
            css_ctr <- ""
            if (center) css_ctr <- "text-align:center; display:inline-block; width:100%;"
            cap_txt <- paste0("<span style=\"color:", col, "; ", css_ctr, "\">Figure ", i, ": ", text , "</span>")
            anchor <- paste0("<a name=\"", refName, "\"></a>")
            if (inline) {
                paste0(anchor, cap_txt)    
            } else {
                list(anchor=anchor, cap_txt=cap_txt)
            }
        },
        
        ref=function(refName, link=FALSE, checkRef=TRUE) {
            
            ## This function puts in a cross reference to a caption. You refer to the
            ## caption with the refName that was passed to fig$cap() (not the code chunk name).
            ## The cross reference can be hyperlinked.
            
            if (checkRef && !refName %in% names(ref)) stop(paste0("fig$ref() error: ", refName, " not found"))
            if (link) {
                paste0("<A HREF=\"#", refName, "\">Figure ", ref[[refName]], "</A>")
            } else {
                paste0("Figure ", ref[[refName]])
            }
        },
        
        ref_all=function(){
            ## For debugging
            ref
        })
})

```

```{r setup, echo=FALSE}

## This chunk replaces the default hook for processing plots. It achieves the purposes,
## of laying out auto-numbered captions, but other functionality may be gone.

library(knitr)
knit_hooks$set(plot = function(x, options) {
    sty <- ""
    if (options$fig.align == 'default') {
        sty <- ""
    } else {
        sty <- paste0(" style=\"text-align:", options$fig.align, ";\"")
    }
    
    if (is.list(options$fig.cap)) {
        ## options$fig.cap is a list returned by the function fig$cap()
        str_caption <- options$fig.cap$cap_txt
        str_anchr <- options$fig.cap$anchor
    } else {
        ## options$fig.cap is a character object (hard coded, no anchor)
        str_caption <- options$fig.cap
        str_anchr <- ""
    }
    
    paste('<figure', sty, '>', str_anchr, '<img src="',
        opts_knit$get('base.url'), paste(x, collapse = '.'),
        '"><figcaption>', str_caption, '</figcaption></figure>',
        sep = '')
    
})
```


```{r read_fig_cap, echo=FALSE}
## This chucnk will read through *this* Rmd file, and attempt to extract all of the 
## labels (not caption text) used for Figure captions. These labels are used
## as anchors, so scanning through the document now will allow us to create cross references
## before the caption actually appears. 

## Get the name of this Rmd file
rmdFn <- knitr::current_input()  # filename of input document

## Read lines and close connection
rmdCon <- file(rmdFn, open = "r")
rmdLines <- readLines(rmdCon)
close(rmdCon)

## Pull out all occurences of at least one back tick, followed 
## by any number of characters, followed by fig$cap (all on one line)
figscap_idx <- grep("`+(.*)fig\\$cap", rmdLines)
rmdLines <- rmdLines[figscap_idx]

## Get rid of everything up until the start of the caption label
## This presumes the caption label is the first argument of fig$cap()
## E.g., fig.cap = fig$cap("my_label", ...)
rmdLinesSansPre <- sub("(.*)fig\\$cap(.*?)[\"']", "", rmdLines)

## Identify everything up until the first quote
match_data <- regexpr("(.*?)[\"']", rmdLinesSansPre)

## Reduce the length by one, because we're not interested in the final quote
attr(match_data, "match.length") <- attr(match_data, "match.length") - 1

## Extract
fig_labels <- regmatches(rmdLinesSansPre, match_data, invert=FALSE)

if (length(fig_labels) > 0) {

    ## Test for duplicates
    if (anyDuplicated(fig_labels) > 0) stop("Duplicate caption labels detected")
    
    ## Create a named list of Figure numbers
    ref <- as.list(1:length(fig_labels))
    names(ref) <- fig_labels
}    

```




### **TASK 1**


The following code sample shows how to import the data needed for the assignement hosted on GitHub: https://raw.githubusercontent.com/juliuspeturg/Assignment2/.
For Task 1 we wrote function with thee inputs y,d,l where y is the input data, d is the number of differenceiation and l is the number of lag´s to plot. the output of the function is 4 plot two ACF´s, PACF and a sprectrum plot of the data.

```{r chunk1,eval=T}
suppressPackageStartupMessages(library(RCurl))
suppressPackageStartupMessages(library(ggplot2))
x = getURL("https://raw.githubusercontent.com/juliuspeturg/Assignment2/master/VEKS.csv")
VEKS = read.csv(text=x,stringsAsFactors = FALSE,header=TRUE)

attach(VEKS)
source("Multiplot.R")

xx <- qplot((1:length(HC.c)),HC.c,geom="line")+
    xlab("Time")+
    ylab("Heat Consumption")+
    ggtitle("Heat Consumption")
yy <- qplot((1:length(Ta.c)),Ta.c,geom="line")+
    xlab("Time")+
    ylab("Air Temperature")+
    ggtitle("Air Temperature")
multiplot(xx,yy,cols = 1)

```
`r fig$cap("heatVStemp", "skrifa", inline=TRUE)`

We notice that there are fluctuations in the heat consumption data on a small scale (day by day and week by week) and also a seasonal fluctuation, indicating more heat consumption during the winter and less during the summer when the ambient air temperature is higher.

```{r chunk1-1,eval=T}
acf.plot <- function(y,d,l){
  source("Multiplot.R")
  suppressPackageStartupMessages(require(ggplot2))
  suppressPackageStartupMessages(require(ggfortify))
  if(d!=0){
    y <- diff(y,difference=d)
  }
  else{}
  #----------Autocovariance--------#
  acf <- acf(y,type="covariance", na.action = na.pass ,plot = FALSE,lag.max=l)
  acfdf <- with(acf, data.frame(lag, acf))
  
  p1 <- ggplot(data = acfdf, mapping = aes(x = lag, y = acf)) +
    geom_hline(aes(yintercept = 0)) +
    geom_segment(mapping = aes(xend = lag, yend = 0))+
    xlab("lag")+
    ylab("ACF")+
    ggtitle("AutoCovariance")
  #----------AutoCorrelation--------#
  c_acf <- acf(y,type="correlation", na.action = na.pass ,plot = FALSE,lag.max=l)
  c_acfdf <- with(c_acf, data.frame(lag, acf))
  ci <- qnorm((1 + 0.95)/2)/sqrt(c_acf$n.used)
  p2 <- ggplot(data = c_acfdf, mapping = aes(x = lag, y = acf)) +
    geom_hline(aes(yintercept = 0)) +
    geom_segment(mapping = aes(xend = lag, yend = 0))+
    ggplot2::geom_hline(yintercept=c(-ci, ci), colour="blue", linetype="dashed")+
    xlab("lag")+
    ylab("ACF")+
    ggtitle("AutoCorrelation")
  
  #----------PACF--------#
  bpacf <- pacf(y,na.action = na.pass,plot = FALSE,lag.max=l)
  bpacfdf <- with(bpacf, data.frame(lag, acf))
  
  p3 <- ggplot(data = bpacfdf, mapping = aes(x = lag, y = acf)) +
    geom_hline(aes(yintercept = 0)) +
    geom_segment(mapping = aes(xend = lag, yend = 0))+
    ggplot2::geom_hline(yintercept=c(-ci, ci), colour="blue", linetype="dashed")+
    xlab("lag")+
    ylab("PACF")+
    ggtitle("Patial AutoCorrelation")
  #----------Spectrum--------#
  k = kernel("daniell", c(20,20))
  specvalues = spec.pgram(na.omit(y), k, taper=0, log = "no",plot = F)
  peak=specvalues$freq[which.max(specvalues$spec)]
  p4 <- autoplot(spec.pgram(na.omit(y), k, taper=0, log = "no",plot = FALSE))+
    geom_vline(xintercept = peak,colour = "red", linetype = "dotted")+
    ggtitle("Spectrum Plot")
  
  multiplot(p1,p2,p3,p4,cols = 2)
}

```

So now we ca use ouer function to analize the heat consumption time series.

```{r chunk2}
acf.plot(VEKS$HC.c,0,50)
```
`r fig$cap("plot1", "Plot of the two ACF´s (covariance and correlation) aswell as the PACF and Sprectrum plot with diff set at zero", inline=TRUE)`


As we can see in `r fig$ref("plot1",link=F)` the ACF tapers of very slowly, this indicats tha the timeseries is nonstationari and we shoud diffirenciat to be able to extract any usueful information from tha plots so next we estimate the difference at lag=1 and take anther look at the outputs.


```{r chunk3}
acf.plot(VEKS$HC.c,1,50)
```
`r fig$cap("plot2", "Plot of the two ACF´s (covariance and correlation) aswell as the PACF and Sprectrum plot with diff set at 1", inline=TRUE)`


In `r fig$ref("plot2",link=F)` We can se a pattern emerge, we see some intresting spikes in lag 12 and lag 24 lag 36 and so on.Both the ACF and PACF show sem cyclical behavior but the largest spike in lag 24. If we look at the spectrum plot of the series we se two destinct spices with frequency of around 0.4 and 0.08. the inverse of these frequencyc is the periog of 24hours and 12hours. 

### **TASK 2**

Intuitively we know that the data should have seasonality within every 24 hours, every week and also within every year.
When predicting for such a short time period (1 hour and 6 hours), we conclude that the 24 hour cycle is the most important seasonality cycle to look at, and therefore we change the data into a time series with frequency 24:

```{r Task2_1}

suppressPackageStartupMessages(require(forecast))
heat_con <- ts(na.omit(HC.c), frequency = 24)  

```

lets examine the forescast for 6 hour ahead
Let us first try a linear model using tslm()

```{r Task2_2}

fit_lm <- tslm(heat_con ~ trend + season)
accuracy(fit_lm)
fcast_lm <- forecast(fit_lm, h=6)
#summary(fit_lm)
autoplot(fcast_lm, include = 168)

```
`r fig$cap("LMforecast", "skrifa um forecast", inline=TRUE)`


After taking the first difference of the heat consumption time series, we inspect the ACF and PACF plots
and estimate that the heat consumption can be described with a ARIMA(3,1,1) process.
Using arima to predict with ARIMA(3,1,1):

```{r Task2_3}

fit1.arima <- arima(na.omit(HC.c), order=c(3,1,1))
#summary(fit1.arima)
accuracy(fit1.arima)
fcast1 <- forecast(fit1.arima,6)
autoplot(fcast1, include = 168)

```
`r fig$cap("ARIMAforecast", "skrifa um arima", inline=TRUE)`

Now we use the auto.arima() function from the forecast package in R. That gives a different model
than we estimated earlier, or ARIMA(4,1,4).

```{r Task2_4}

# auto.arima(na.omit(HC.c))  
# Gives ARIMA(4,1,4)
# Let's now use this input to predict:
fit2.arima <- arima(na.omit(HC.c), order=c(4,1,4))
#summary(fit2.arima)
accuracy(fit2.arima)
fcast2 <- forecast(fit2.arima,6)
autoplot(fcast2, include = 168)

```
`r fig$cap("autoarima(4,1,4)forecast", "skrifa um arima", inline=TRUE)`

lets examine the forescast for 1 hour ahead with the same methods as above



### **TASK 3**

For this task we did simmilar things as in task 1, made a function to plot the cross correlation of tha possible meshurments that might be used to aid in forecasting the the head consumption. The meshurements for air temperatur, wind speed ang global radiation were used. and tha cross correlation of the three meshurements and the heat consumtion were evaluated as can be seen below.

```{r chunk4,eval=T}
CCF.plot <- function(d,l){
  source("Multiplot.R")
  suppressPackageStartupMessages(require(forecast))  
  suppressPackageStartupMessages(require(ggplot2))
  suppressPackageStartupMessages(require(ggfortify))
  
  HeatConsumption <- VEKS$HC.c
  Air_Temperature <- VEKS$Ta.c
  WindSpeed       <- VEKS$W.c 
  GlobalRadiation <- VEKS$GR.c
  
  if(d!=0){
    HeatConsumption <- diff(VEKS$HC.c,difference=d)
    Air_Temperature     <- diff(VEKS$Ta.c,difference=d)
    WindSpeed       <- diff(VEKS$W.c, difference=d) 
    GlobalRadiation <- diff(VEKS$GR.c,difference=d)
  }
  else{}
  
  ccf1 <- autoplot(stats::ccf(HeatConsumption,Air_Temperature,plot=FALSE,na.action=na.pass,lag.max = l))+
    geom_vline(xintercept = 0,colour = "red", linetype = "dotted")+
    ggtitle("Air temp. & Heat Consumption")


  ccf2 <- autoplot(stats::ccf(HeatConsumption,GlobalRadiation,plot=FALSE,na.action=na.pass,lag.max = l))+
    geom_vline(xintercept = 0,colour = "red", linetype = "dotted")+
    ggtitle("Global rad. & Heat Consumption")

  
  ccf3 <- autoplot(stats::ccf(HeatConsumption,WindSpeed,plot=FALSE,na.action=na.pass,lag.max = l))+
    geom_vline(xintercept = 0,colour = "red", linetype = "dotted")+
    ggtitle("Wind & Heat Consumption")

  
  multiplot(ccf1,ccf2,ccf3,cols=1)
}

```

the input to the plotfunction is  d and l for the lag.max, the difference to be used on the data. so lets see what tha cross correlation looks like at zero difference

```{r chunk5}
CCF.plot(0,20)
```
`r fig$cap("plot3", "Plot of the cross correlation between the consumtion and Air Air_Temperature(top),Windspeed (middle) and Global Radiation(bottom) with diff set at zero. The red dottet line is Lag=0", inline=TRUE)`

As wee can see in `r fig$ref("plot3")` there is not much information to be extracted from these plot. agen due to the series being non-stationari so we replot again with now with diff=1.

```{r chunk6}
CCF.plot(1,20)
```
`r fig$cap("plot4", "Plot of the cross correlation between the consumtion and Air Air_Temperature(top),Windspeed (middle) and Global Radiation(bottom) with diff set at 1. The red dottet line is Lag=0", inline=TRUE)`

From `f fig$ref("plot4")` we think therea might be some indicator of lagged effect in lag 0 and lag 1 for the air temperature and windspeed and some at lag 1 for global radiation so now it shoud be interesting what pre-whitening will do to tha data. lets start with looking at air temperature

```{r chunk7,eval=F,echo=F}
HeatConsumption <- VEKS$HC.c
Air_Temperature <- VEKS$Ta.c
WindSpeed       <- VEKS$W.c 
GlobalRadiation <- VEKS$GR.c

x1 <- ggAcf(Air_Temperature,na.action=na.pass)+ggtitle("Acf for Air Temp. diff 0")
x2 <- ggPacf(Air_Temperature,na.action=na.pass)+ggtitle("Pacf for Air Temp. diff 0")
x3 <- ggAcf(diff(Air_Temperature,difference=1),na.action=na.pass)+ggtitle("Acf for Air Temp. diff 1")
x4 <- ggPacf(diff(Air_Temperature,difference=1),na.action=na.pass)+ggtitle("Pacf for Air Temp. diff 1")

multiplot(x1,x3,x2,x4,cols=2)
```

```{r chunk8,eval=T,echo=F}
pw.air_temp<-auto.arima(VEKS$Ta.c)
#pw.temp = ARIMA(3,1,2) with drift
pw.Hc<-residuals(arima(VEKS$HC.c,order=c(3,1,2),fixed=c(pw.air_temp$coef[1:5])))
x <- pw.air_temp$residuals
y <- pw.Hc
ggCcf(x,y,lag.max=25)+ggtitle("Pre-whitening series: Air temp. & Heat consumption")+
    geom_vline(xintercept = 0,colour = "red", linetype = "dotted")+
    geom_vline(xintercept = c(-12,12),colour = "red", linetype = "dotted")+
    geom_vline(xintercept = c(-24,24),colour = "red", linetype = "dotted")
```
`r fig$cap("plot5", "Pre-whitening. here we can strong indicator in lag 1", inline=TRUE)`


```{r chunk9,eval=T,echo=F}
pw.wind<-auto.arima(VEKS$W.c)
#pw.wind = ARIMA(4,1,2)
pw.Hc<-residuals(arima(VEKS$HC.c,order=c(4,1,2),fixed=c(pw.wind$coef)))
x <- pw.wind$residuals
y <- pw.Hc
ggCcf(x,y,lag.max=25)+ggtitle("Pre-whitening series: Wind & Heat consumption")+
    geom_vline(xintercept = 0,colour = "red", linetype = "dotted")+
    geom_vline(xintercept = c(-12,12),colour = "red", linetype = "dotted")+
    geom_vline(xintercept = c(-24,24),colour = "red", linetype = "dotted")
```
`r fig$cap("plot6", "Pre-whitening. here we can strong indicator in lag 1", inline=TRUE)`


```{r chunk10,eval=T,echo=F}
pw.radiation<-auto.arima(VEKS$GR.c)
#pw.radiation = ARIMA(3,1,2)
pw.Hc<-residuals(arima(VEKS$HC.c,order=c(3,1,2),fixed=c(pw.radiation$coef)))
x <- pw.radiation$residuals
y <- pw.Hc
ggCcf(x,y,lag.max=25)+ggtitle("Pre-whitening series: radiation & Heat consumption")+
    geom_vline(xintercept = 0,colour = "red", linetype = "dotted")+
    geom_vline(xintercept = c(-12,12),colour = "red", linetype = "dotted")+
    geom_vline(xintercept = c(-24,24),colour = "red", linetype = "dotted")
```
`r fig$cap("plot7", "Pre-whitening. here we can strong indicator in lag 1", inline=TRUE)`
