---
title: "Project 1"
author: 
- "Júlíus Pétur Guðjohnsen"
- "Sigurður Jónsson"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output: 
  html_document: 
    fig_height: 4
    fig_width: 7
    fig_caption: yes
    self_contained: yes
    smart: no
---

```{r auto_num_functions, echo=FALSE}
# A function for generating captions and cross-references

fig <- local({
    i <- 0
    list(
        cap=function(refName, text, center=FALSE, col="black", inline=FALSE) {
            i <<- i + 1
            ref[[refName]] <<- i
            css_ctr <- ""
            if (center) css_ctr <- "text-align:center; display:inline-block; width:100%;"
            cap_txt <- paste0("<span style=\"color:", col, "; ", css_ctr, "\">Figure ", i, ": ", text , "</span>")
            anchor <- paste0("<a name=\"", refName, "\"></a>")
            if (inline) {
                paste0(anchor, cap_txt)    
            } else {
                list(anchor=anchor, cap_txt=cap_txt)
            }
        },
        
        ref=function(refName, link=FALSE, checkRef=TRUE) {
            
            ## This function puts in a cross reference to a caption. You refer to the
            ## caption with the refName that was passed to fig$cap() (not the code chunk name).
            ## The cross reference can be hyperlinked.
            
            if (checkRef && !refName %in% names(ref)) stop(paste0("fig$ref() error: ", refName, " not found"))
            if (link) {
                paste0("<A HREF=\"#", refName, "\">Figure ", ref[[refName]], "</A>")
            } else {
                paste0("Figure ", ref[[refName]])
            }
        },
        
        ref_all=function(){
            ## For debugging
            ref
        })
})

```

```{r setup, echo=FALSE}

## This chunk replaces the default hook for processing plots. It achieves the purposes,
## of laying out auto-numbered captions, but other functionality may be gone.

library(knitr)
knit_hooks$set(plot = function(x, options) {
    sty <- ""
    if (options$fig.align == 'default') {
        sty <- ""
    } else {
        sty <- paste0(" style=\"text-align:", options$fig.align, ";\"")
    }
    
    if (is.list(options$fig.cap)) {
        ## options$fig.cap is a list returned by the function fig$cap()
        str_caption <- options$fig.cap$cap_txt
        str_anchr <- options$fig.cap$anchor
    } else {
        ## options$fig.cap is a character object (hard coded, no anchor)
        str_caption <- options$fig.cap
        str_anchr <- ""
    }
    
    paste('<figure', sty, '>', str_anchr, '<img src="',
        opts_knit$get('base.url'), paste(x, collapse = '.'),
        '"><figcaption>', str_caption, '</figcaption></figure>',
        sep = '')
    
})
```


```{r read_fig_cap, echo=FALSE}
## This chucnk will read through *this* Rmd file, and attempt to extract all of the 
## labels (not caption text) used for Figure captions. These labels are used
## as anchors, so scanning through the document now will allow us to create cross references
## before the caption actually appears. 

## Get the name of this Rmd file
rmdFn <- knitr::current_input()  # filename of input document

## Read lines and close connection
rmdCon <- file(rmdFn, open = "r")
rmdLines <- readLines(rmdCon)
close(rmdCon)

## Pull out all occurences of at least one back tick, followed 
## by any number of characters, followed by fig$cap (all on one line)
figscap_idx <- grep("`+(.*)fig\\$cap", rmdLines)
rmdLines <- rmdLines[figscap_idx]

## Get rid of everything up until the start of the caption label
## This presumes the caption label is the first argument of fig$cap()
## E.g., fig.cap = fig$cap("my_label", ...)
rmdLinesSansPre <- sub("(.*)fig\\$cap(.*?)[\"']", "", rmdLines)

## Identify everything up until the first quote
match_data <- regexpr("(.*?)[\"']", rmdLinesSansPre)

## Reduce the length by one, because we're not interested in the final quote
attr(match_data, "match.length") <- attr(match_data, "match.length") - 1

## Extract
fig_labels <- regmatches(rmdLinesSansPre, match_data, invert=FALSE)

if (length(fig_labels) > 0) {

    ## Test for duplicates
    if (anyDuplicated(fig_labels) > 0) stop("Duplicate caption labels detected")
    
    ## Create a named list of Figure numbers
    ref <- as.list(1:length(fig_labels))
    names(ref) <- fig_labels
}    

```




### **Assignment 1**
#### Question 1


<!-- Hérna er cóðinn fyrir fyrsta lið án evaluation -->
The following code sample shows how to import the data needed for the assignement hosted on a https server "https://raw.githubusercontent.com/juliuspeturg/Assignment1/master/fuel.csv".
To better understand the data a summary of the imported data is displayed bellow aswell as a lineplot of the fuel price index, see  `r fig$ref("fuel_data_plot",link=TRUE)`.


<!--$$x=\frac{1+y}{1+2z^2}$$
$$\begin{equation} x=y+3 \label{eq:xdef}
\end{equation}$$
In equation (\ref{eq:xdef}) we saw $\dots$ -->


```{r chunk1}
suppressPackageStartupMessages(library(RCurl))
# load csv file from URL
x <- getURL("https://raw.githubusercontent.com/juliuspeturg/Assignment1/master/fuel.csv")
fuel <- read.csv(text=x,stringsAsFactors = FALSE)
summary(fuel)
attach(fuel)

windows.options(width=7, height=4)
op <- par(cex.lab=0.8, cex.axis=0.8, cex.main=0.8, cex.sub=0.8)
par(mar=c(4.1,4.1,1.1,1.1))
par(mfrow=c(1,1))
plot(fpi~rtime,type = "l",main="",ylab = "",xlab = "")
grid()
title(main="Fuel price index in the US", line = 0.5)
title(xlab="Time", line = 2)
title(ylab="Fpi", line = 2)

```
`r fig$cap("fuel_data_plot", "The fuel price index from 1979 to 2004", inline=TRUE)`


#### Question 2
For statinary time series the ACF will drop to zero very quickly however for nonstationary time series the ACF will taper off slowly. As we can see by looking at the ACF of the fuel price index the ACF tapers off slowly, from that we can conclude that the timeseries is nonstationary and therefor the mean of the series is not going to informe us a lot on the timeseries, see figure below for ACF of timeseries  
```{r eval=TRUE, echo=TRUE,message=FALSE}
#Generate timeseries object for fuelindex
fuel.ts = ts(data = fuel$fpi, start = c(1979,1),frequency = 12)
#evaluate the mean of the timeseries
# #cat(c("mean_fpi:",mean(fpi)))
op <- par(cex.lab=0.8, cex.axis=0.8, cex.main=0.8, cex.sub=0.8)
par(mar=c(4.1,4.1,1.1,1.1))
par(mfrow=c(1,1))
acf(fuel.ts,main="",ylab = "",xlab = "")
grid()
title(main="ACF for fuel price index", line = 0.5)
title(xlab="Lag", line = 2)
title(ylab="ACF", line = 2)

```
`r fig$cap("fuel_ACF", "Here we se tha ACF for the fuel price index, as we can see the ACF tapers of", inline=TRUE)`

#### Question 3

```{r eval=TRUE, echo=TRUE,message=FALSE}
# Linear Model
mod <- lm(fpi~rtime)
summary(mod)
```

#### Question 4

```{r eval=FALSE, echo=TRUE,message=FALSE,results=FALSE}
# Trainingset
myts <- ts(fpi, start = c(1979,1), end =c(2003,12),frequency = 12)
# Testset
myts2 <- ts(fpi, start = c(2004,1), end =c(2004,12),frequency = 12)
require(forecast)
op <- par(cex.lab=0.8, cex.axis=0.8, cex.main=0.8, cex.sub=0.8)
par(mar=c(4.1,4.1,1.1,1.1))
par(mfrow=c(1,1))

monthplot(myts,main="",ylab = "",xlab = "")
grid()
title(main="Fuel price index in the US", line = 0.5)
title(xlab="Time", line = 2)
title(ylab="Fpi", line = 2)

seasonplot(myts,main="",ylab = "",xlab = "")
grid()
title(main="Seasonal Plot", line = 0.5)
title(xlab="Month", line = 2)
title(ylab="Factor", line = 2)

fit1 <- stl(myts,s.window = "period")
#3 eða 4
plot(fit1,main="")
grid()
title(main="Multiplot", line = 0.5)


```


```{r eval=TRUE, echo=FALSE,message=FALSE}
myts <- ts(fpi, start = c(1979,1), end =c(2003,12),frequency = 12)
# Testset
myts2 <- ts(fpi, start = c(2004,1), end =c(2004,12),frequency = 12)
require(forecast)
op <- par(cex.lab=0.8, cex.axis=0.8, cex.main=0.8, cex.sub=0.8)
par(mar=c(4.1,4.1,1.1,1.1))
par(mfrow=c(1,1))
monthplot(myts,main="",ylab = "",xlab = "")
grid()
title(main="Fuel price index in the US", line = 0.5)
title(xlab="Time", line = 2)
title(ylab="Fpi", line = 2)
```
`r fig$cap("fuel_monthplot", "Monthplot from tha Forecast package", inline=TRUE)`
```{r eval=TRUE, echo=FALSE,message=FALSE}
op <- par(cex.lab=0.8, cex.axis=0.8, cex.main=0.8, cex.sub=0.8)
par(mar=c(4.1,4.1,1.1,1.1))
par(mfrow=c(1,1))
seasonplot(myts,main="",ylab = "",xlab = "")
grid()
title(main="Seasonal Plot", line = 0.5)
title(xlab="Month", line = 2)
title(ylab="Factor", line = 2)
```
`r fig$cap("fuel_seasonal", "Seasonal plot from the Forecast package", inline=TRUE)`

```{r eval=TRUE, echo=FALSE,message=FALSE}
op <- par(cex.lab=0.8, cex.axis=0.8, cex.main=0.8, cex.sub=0.8)
par(mar=c(4.1,4.1,1.1,1.1))
par(mfrow=c(1,1))
fit1 <- stl(myts,s.window = "period")
#3 eða 4
plot(fit1,main="")
title(main="Multiplot", line = 2)
```
`r fig$cap("fuel_seasonal_total", "Seasonal plot of the data", inline=TRUE)`


#### Question 5

```{r eval=TRUE, echo=TRUE,message=FALSE}
fuel.ts <- ts(fuel$fpi, frequency=12, start=c(1979,1))
fuel_train <- window(fuel.ts, start=c(1979,1), end=c(2003,12))
fuel_test <- window(fuel.ts, start=c(2004,1), end=c(2004,12))
```

#### Question 6

```{r eval=TRUE, echo=TRUE,message=FALSE}
fit2 <- ses(fuel_train,h = 12, alpha = 0.1,exponential=TRUE)
f.fit2 <- forecast(fit2,12)
op <- par(cex.lab=0.8, cex.axis=0.8, cex.main=0.8, cex.sub=0.8)
par(mar=c(4.1,4.1,1.1,1.1))
par(mfrow=c(1,1))
plot(f.fit2,main="",ylab="",xlab="")
lines(fuel_train)
lines(fuel_test,col=2)
grid()
title(main="Simple Explonential Smoothing Forecast", line = 0.5)
title(xlab="Time", line = 2)
title(ylab="Fpi", line = 2)
```
`r fig$cap("SES_fuel", "Simple exponential smoothing forecast for the year 2004", inline=TRUE)`

#### Question 7

```{r eval=TRUE, echo=TRUE,message=FALSE}
fit3 <- ses(fuel_train,h = 12, exponential=FALSE)
f.fit3 <- forecast(fit3,12)
op <- par(cex.lab=0.8, cex.axis=0.8, cex.main=0.8, cex.sub=0.8)
par(mar=c(4.1,4.1,1.1,1.1))
par(mfrow=c(1,1))
plot(f.fit3,main="",ylab="",xlab="")
lines(fuel_train)
lines(fuel_test,col=2)
grid()
title(main="Local Trend Forecast", line = 0.5)
title(xlab="Time", line = 2)
title(ylab="Fpi", line = 2)
```
`r fig$cap("local_fuel", "Local trend forecast for the year 2004", inline=TRUE)`

#### Question 8

```{r eval=TRUE, echo=TRUE,message=FALSE}
fit4 <- HoltWinters(myts, beta = FALSE,gamma = FALSE)
f.fit4 <- forecast(fit4,12)
op <- par(cex.lab=0.8, cex.axis=0.8, cex.main=0.8, cex.sub=0.8)
par(mar=c(4.1,4.1,1.1,1.1))
par(mfrow=c(1,1))
plot(f.fit4,main="",ylab="",xlab="")
lines(fuel_train)
lines(fuel_test,col=2)
grid()
title(main="HoltWinters Forecast", line = 0.5)
title(xlab="Time", line = 2)
title(ylab="Fpi", line = 2)
```
`r fig$cap("holtwinters_fuel", "HoltWinters forecast for the year 2004", inline=TRUE)`

#### Question 9

```{r eval=TRUE, echo=TRUE,message=FALSE}
fit5 <- rwf(myts, h=12, drift=FALSE, fan=FALSE, lambda=NULL, biasadj=FALSE)
f.fit5 <- forecast(fit5,12)
op <- par(cex.lab=0.8, cex.axis=0.8, cex.main=0.8, cex.sub=0.8)
par(mar=c(4.1,4.1,1.1,1.1))
par(mfrow=c(1,1))
plot(f.fit5,main="",ylab="",xlab="")
lines(fuel_train)
lines(fuel_test,col=2)
grid()
title(main="Random Walk Forecast", line = 0.5)
title(xlab="Time", line = 2)
title(ylab="Fpi", line = 2)
```
`r fig$cap("randomwalk_fuel", "Random walk forecast for the year 2004 without drift", inline=TRUE)`

#### Question 10

```{r eval=TRUE, echo=TRUE,message=FALSE}
RMSE_SES <- accuracy(f.fit2)
RMSE_LT <- accuracy(f.fit3)
RMSE_HW <- accuracy(f.fit4)
RMSE_RW <- accuracy(f.fit5)

ACC <- vector(length=4); ACC[1]=RMSE_SES[2]; ACC[2]=RMSE_LT[2]; ACC[3]=RMSE_HW[2]; ACC[4]=RMSE_RW[2]

op <- par(cex.lab=0.8, cex.axis=0.8, cex.main=0.8, cex.sub=0.8)
par(mar=c(4.1,4.1,1.1,1.1))
par(mfrow=c(1,1))
barplot(ACC,main="",names.arg=c("SES", "LT", "HW","RW"),xlab="")
title(main="Error of Forecasts", line = 0.5)
title(ylab="RMSE (%)", line = 2)
```
`r fig$cap("Accuracy", "Barplot showing the Error of each of the methods", inline=TRUE)`


---

### **Assignment 2**
#### Question 1

```{r eval=TRUE, echo=TRUE,message=FALSE}
library(RCurl)
x <- getURL("https://raw.githubusercontent.com/juliuspeturg/Assignment1/master/simdata.csv")
simdata <- read.csv(text=x,stringsAsFactors = FALSE,header=TRUE, sep="")

# A2-Q1: Plot the data

attach(simdata)

# á sama grafi
plot(y1, type="l",col='red', main="Simdata time series",
     xlab="Time", ylab="Data")
lines(y2,col='blue')
lines(y3,col='green')
legend('topright', legend=c("y1", "y2","y3"),
       col=c('red','blue','green'), lty=1, bty='y', cex=.85)
```
`r fig$cap("simdata_plot", "Plot of the three simdata series", inline=TRUE)`

#### Question 2

```{r eval=TRUE, echo=TRUE,message=FALSE}
op <- par(cex.lab=0.8, cex.axis=0.8, cex.main=0.8, cex.sub=0.8)
par(mar=c(4.1,4.1,1.1,1.1))
par(mfrow=c(3,2))

# For cycle to plot the ACF and PACF for all the column in simdata
for(i in 1:ncol(simdata)){
  acf(simdata[[i]],main="",ylab="",xlab="")
  title(main=paste("Series ",colnames(simdata[i])), line = 0.5)
  title(xlab="Lag", line = 2)
  title(ylab="ACF", line = 2)
  
  pacf(simdata[[i]],main="",ylab="",xlab="")
  title(main=paste("Series ",colnames(simdata[i])), line = 0.5)
  title(xlab="Lag", line = 2)
  title(ylab="PACF", line = 2)
}

```
`r fig$cap("simdata_ACF_PACF", "Plot of the ACF and PACF for all three series", inline=TRUE)`

#### Question 3

```{r eval=TRUE, echo=TRUE,message=FALSE}
# A2-Q3: Identify the models using arima  <<< ATH BETUR
arima(y1)
arima(y2)
arima(y3)
auto.arima(y1) # líka hægt að nota auto.arima
```


#### Question 4

```{r eval=TRUE, echo=TRUE,message=FALSE}
ts.sim <- arima.sim(list(order = c(1,0,0), ar = 0.9), n = 250)
```

#### Question 5

```{r eval=TRUE, echo=TRUE,message=FALSE}
layout(matrix(c(1,1,2,3), 2, 2, byrow = TRUE))  
plot(ts.sim, main="Simulated time series ts.sim")
acf(ts.sim, main="ACF for ts.sim")
pacf(ts.sim, main="PACF for ts.sim")
```
`r fig$cap("Sim_ARIMA", "Simulated Arima series", inline=TRUE)`
#### Question 6

```{r eval=TRUE, echo=TRUE,message=FALSE}
acf.plot <- function(y){
  layout(matrix(c(1,1,2,3), 2, 2, byrow = TRUE)) 
  plot(y, main=c("Simulated time series",deparse(substitute(y))))
  acf(y, main=c("ACF for",deparse(substitute(y))))
  pacf(y,main=c("PACF for",deparse(substitute(y))))
}
```

#### Question 7

```{r eval=TRUE, echo=TRUE,message=FALSE,fig.show='hold'}
ts.sim1 <- arima.sim(n=250, list(ar=c(0.5)))        
ts.sim2 <- arima.sim(n=250, list(ar=c(-0.5)))
ts.sim3 <- arima.sim(n=250, list(ar=c(0.5,0.2)))
ts.sim4 <- arima.sim(n=250, list(ar=c(0.7)))
ts.sim5 <- arima.sim(n=250, list(ar=c(0.5,-0.2)))
ts.sim6 <- arima.sim(n=250, list(ma=c(0.5,0.5)))
ts.sim7 <- arima.sim(n=250, list(ma=c(-0.9)))
ts.sim8 <- arima.sim(n=250, list(ar=c(0.5), ma=c(-0.8)))

# Plotting the simulated series with the acf function
acf.plot(ts.sim1)
acf.plot(ts.sim2)
acf.plot(ts.sim3)
acf.plot(ts.sim4)
acf.plot(ts.sim5)
acf.plot(ts.sim6)
acf.plot(ts.sim7)
acf.plot(ts.sim8)
```
`r fig$cap("ACF_ALL", "Plot of the simulated series, for all the series í question 7 ", inline=TRUE)`

#### Question 8

```{r eval=TRUE, echo=TRUE,message=FALSE}
ts.sim9 <- arima.sim(list(order = c(1,1,1), ar = 0.7, ma=0.3), n = 250)

ts.sim9_diff1 <- diff(ts.sim9, differences=1) #first difference
plot.ts(ts.sim9, col="red")
lines(ts.sim9_diff1, col="blue")
legend('topleft', legend=c("ts.sim9", "ts.sim9_diff1"),
       col=c('red','blue'), lty=1, bty='y', cex=.85)
grid()
```
`r fig$cap("series_and_simseries", "some fucked up shit ", inline=TRUE)`

---

### **Assignment 3**
#### Question 1

```{r eval=TRUE, echo=TRUE,message=FALSE}
x <- getURL("https://vincentarelbundock.github.io/Rdatasets/csv/datasets/AirPassengers.csv")
flightdata <- read.csv(text=x,stringsAsFactors = FALSE,header=TRUE)
flightdata$X <- NULL
y <- flightdata$AirPassengers

SimpleMovingAverage <- function(y,n){
  #this function takes in a dataframe having column 1 as data and
  # column 2 the values of the time series
  FC=matrix(nrow=(nrow(y)-(n-1)),ncol=2)
  Y = y[[2]]
  for(i in 1:(nrow(y)-(n-1))){
  FC[i,2] = sum(Y[1:n])/n
  FC[i,1] = y[(n+i),1]
  Y <- Y[-1] 
  }
  FC <- na.omit(FC)
  FC
}

A <- SimpleMovingAverage(flightdata,10)
B <- SimpleMovingAverage(flightdata,5)


plot(flightdata$time,flightdata$AirPassengers,type="l",col="blue")
lines(A[,1],A[,2],col="red")
lines(B[,1],B[,2],col="green")
grid()
```
`r fig$cap("flightdata_sam", "Simple moving average for flight data", inline=TRUE)`

#### Questin 2

```{r eval=TRUE, echo=TRUE,message=FALSE}
ExponentialSmoothing <- function(y,alpha){
  actual <- y[[2]]
  forcast <- matrix(nrow=length(actual),ncol=2)
  forcast[1,2] <- actual[1]
  for(i in 2:length(actual)){
    forcast[i,2] <- alpha*actual[i-1] + (1-alpha)*forcast[i-1,2]
  }
  forcast[,1] <- y[[1]]
  forcast
}

C <- ExponentialSmoothing(flightdata,1)
D <- ExponentialSmoothing(flightdata,0.5)


plot(flightdata$time,flightdata$AirPassengers,type="l",col="blue")
lines(C[,1],C[,2],col="red")
lines(D[,1],D[,2],col="green")
grid()
```
`r fig$cap("flightdata_ses", "simple explonential smoothing", inline=TRUE)`

#### Question 3

```{r eval=TRUE, echo=TRUE,message=FALSE}
DoubleExponentialSmoothing <- function(y,alpha,beta){
  actual <- y[[2]]
  trend <- vector(length=length(actual))
  forcast <- matrix(nrow=length(actual),ncol=2)
  forcast[1,2] <- actual[1]
  trend[1] <- actual[2] - actual[1]
  for(i in 2:length(actual)){
    forcast[i,2] <- alpha*actual[i-1] + (1-alpha)*(forcast[i-1,2]+trend[i-1])
    trend[i]     <- beta*(forcast[i,2]-forcast[i-1,2])+(1-beta)*trend[i-1]
  }
  forcast[,1] <- y[[1]]
  forcast
  return((list(forcast,trend[length(trend)])))
}

E <- DoubleExponentialSmoothing(flightdata,0.2,0.3)
G <- matrix(unlist(E[1]),ncol=2,byrow=FALSE)
plot(flightdata$time,flightdata$AirPassengers,type="l",col="blue")
lines(G[,1],G[,2],col="green")
grid()
```
`r fig$cap("flightdata_des", "double exponential smoothing", inline=TRUE)`

#### Question 4

```{r eval=TRUE, echo=TRUE,message=FALSE}
E <- DoubleExponentialSmoothing(flightdata,0.2,0.3)
G <- matrix(unlist(E[1]),ncol=2,byrow=FALSE)
trend <- as.numeric(E[2])
predicted <- matrix(nrow = 20,ncol = 2)
predicted[1,2] <- G[nrow(G),2]
difference <- G[nrow(G),1]-G[(nrow(G)-1),1]
predicted[1,1] <- G[nrow(G),1]+difference
for(i in 2:nrow(predicted)){
  predicted[i,2] <- predicted[1,2]+trend*(i-1)
  predicted[i,1] <- predicted[i-1,1]+difference
}

stora <- rbind(G,predicted)
plot(stora[,1],stora[,2],col="white",xlim=c(min(stora[,1]),max(stora[,1])),ylim=c(min(stora[,2]),(max(stora[,2])+50)))
lines(flightdata$time,flightdata$AirPassengers,type="l",col="blue")
lines(G[,1],G[,2],col="green")
lines(predicted[,1],predicted[,2],col="red")
grid()
```
`r fig$cap("flightdata_des_forecast", "double exponential smoothing with forecast 20 steps ahead", inline=TRUE)`

---

### **Assignment 4**
#### Question 1

```{r eval=TRUE, echo=TRUE,message=FALSE}
require(googleVis)
op <- options(gvis.plot.tag='chart')

require(RCurl)
x = getURL("https://raw.githubusercontent.com/juliuspeturg/Assignment1/master/gamma-visitolur.csv")
gamma <- read.csv(text=x, stringsAsFactors = FALSE, header=TRUE)

gamma$Date <- as.Date(gamma$Date, "%Y-%m-%d")
gamma <- na.omit(gamma)
names(gamma)[names(gamma)=="GAMMA.v.ed.sit.f6.lur"] <- "GAMMA Equity Index"

```

```{r,results='asis', tidy=FALSE,, fig.width = 7, fig.height = 7}
  A <- gvisAnnotationChart(gamma, datevar = 'Date', numvar = 'GAMMA Equity Index', date.format = "YY-mm-dd")
plot(A)  
```
`r fig$cap("gamma_index", "Plot of the gamma equity index", inline=TRUE)`
```{r eval=TRUE,echo=FALSE}
## Set options back to original options
options(op)
```

#### Question 2

```{r eval=TRUE, echo=TRUE,message=FALSE}
max_row_training = max(which(gamma$Date=="2014-12-31"))
gamma_train <- head(gamma, n=max_row_training)
gamma_test <- tail(gamma, n=(nrow(gamma)-max_row_training))


plot(gamma, type="l", col="white", main="Split of test & training set")
lines(gamma_train, type="l", col=2)
lines(gamma_test, type="l", col=3)
legend('bottomright', legend=c("Training set", "Test set"),
       col=c(2,3), lty=1, bty='y', cex=.85)
grid()
```
`r fig$cap("gamma_train_test", "A plot displaying the splt into trainingset and testset", inline=TRUE)`

#### Question 3

```{r eval=TRUE, echo=TRUE,message=FALSE}
gamma.ts <- ts(gamma$`GAMMA Equity Index`)
gamma_train.ts <- ts(gamma_train$`GAMMA Equity Index`)
gamma_test.ts <- ts(gamma_test$`GAMMA Equity Index`) 

# fit_gamma <- ses(gamma_training$`GAMMA Equity Index`)
require(forecast)
fit_gamma <- HoltWinters(gamma_train.ts, beta = FALSE,gamma = FALSE)
f.fit_gamma <- forecast(fit_gamma, 20)
plot(f.fit_gamma)
grid()
```
`r fig$cap("holtwinters_gamma", "holt winters forecast for the gammaindex", inline=TRUE)`

```{r eval=TRUE, echo=TRUE,message=FALSE}
accuracy(f.fit_gamma)

```

