---
title: "Project 1"
author: 
- "Júlíus Pétur Guðjohnsen"
- "Sigurður Jónsson"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output: html_document
fig_caption: yes
self_contained: yes
smart: no

---

```{r auto_num_functions, echo=FALSE}
# A function for generating captions and cross-references

fig <- local({
    i <- 0
    list(
        cap=function(refName, text, center=FALSE, col="black", inline=FALSE) {
            i <<- i + 1
            ref[[refName]] <<- i
            css_ctr <- ""
            if (center) css_ctr <- "text-align:center; display:inline-block; width:100%;"
            cap_txt <- paste0("<span style=\"color:", col, "; ", css_ctr, "\">Figure ", i, ": ", text , "</span>")
            anchor <- paste0("<a name=\"", refName, "\"></a>")
            if (inline) {
                paste0(anchor, cap_txt)    
            } else {
                list(anchor=anchor, cap_txt=cap_txt)
            }
        },
        
        ref=function(refName, link=FALSE, checkRef=TRUE) {
            
            ## This function puts in a cross reference to a caption. You refer to the
            ## caption with the refName that was passed to fig$cap() (not the code chunk name).
            ## The cross reference can be hyperlinked.
            
            if (checkRef && !refName %in% names(ref)) stop(paste0("fig$ref() error: ", refName, " not found"))
            if (link) {
                paste0("<A HREF=\"#", refName, "\">Figure ", ref[[refName]], "</A>")
            } else {
                paste0("Figure ", ref[[refName]])
            }
        },
        
        ref_all=function(){
            ## For debugging
            ref
        })
})

```

```{r setup, echo=TRUE}

## This chunk replaces the default hook for processing plots. It achieves the purposes,
## of laying out auto-numbered captions, but other functionality may be gone.

library(knitr)
knit_hooks$set(plot = function(x, options) {
    sty <- ""
    if (options$fig.align == 'default') {
        sty <- ""
    } else {
        sty <- paste0(" style=\"text-align:", options$fig.align, ";\"")
    }
    
    if (is.list(options$fig.cap)) {
        ## options$fig.cap is a list returned by the function fig$cap()
        str_caption <- options$fig.cap$cap_txt
        str_anchr <- options$fig.cap$anchor
    } else {
        ## options$fig.cap is a character object (hard coded, no anchor)
        str_caption <- options$fig.cap
        str_anchr <- ""
    }
    
    paste('<figure', sty, '>', str_anchr, '<img src="',
        opts_knit$get('base.url'), paste(x, collapse = '.'),
        '"><figcaption>', str_caption, '</figcaption></figure>',
        sep = '')
    
})
```


```{r read_fig_cap, echo=FALSE}
## This chucnk will read through *this* Rmd file, and attempt to extract all of the 
## labels (not caption text) used for Figure captions. These labels are used
## as anchors, so scanning through the document now will allow us to create cross references
## before the caption actually appears. 

## Get the name of this Rmd file
rmdFn <- knitr::current_input()  # filename of input document

## Read lines and close connection
rmdCon <- file(rmdFn, open = "r")
rmdLines <- readLines(rmdCon)
close(rmdCon)

## Pull out all occurences of at least one back tick, followed 
## by any number of characters, followed by fig$cap (all on one line)
figscap_idx <- grep("`+(.*)fig\\$cap", rmdLines)
rmdLines <- rmdLines[figscap_idx]

## Get rid of everything up until the start of the caption label
## This presumes the caption label is the first argument of fig$cap()
## E.g., fig.cap = fig$cap("my_label", ...)
rmdLinesSansPre <- sub("(.*)fig\\$cap(.*?)[\"']", "", rmdLines)

## Identify everything up until the first quote
match_data <- regexpr("(.*?)[\"']", rmdLinesSansPre)

## Reduce the length by one, because we're not interested in the final quote
attr(match_data, "match.length") <- attr(match_data, "match.length") - 1

## Extract
fig_labels <- regmatches(rmdLinesSansPre, match_data, invert=FALSE)

if (length(fig_labels) > 0) {

    ## Test for duplicates
    if (anyDuplicated(fig_labels) > 0) stop("Duplicate caption labels detected")
    
    ## Create a named list of Figure numbers
    ref <- as.list(1:length(fig_labels))
    names(ref) <- fig_labels
}    

```




### Assignment 1
#### Question 1


<!-- Hérna er cóðinn fyrir fyrsta lið án evaluation -->
The following code sample shows how to import the data needed for the assignement hosted on a https server "https://raw.githubusercontent.com/juliuspeturg/Assignment1/master/fuel.csv".
To better understand the data a summary of the imported data is displayed bellow aswell as a lineplot of the fuel price index

down vote
accepted
I think you should be able to use regular html comments:

<!--$$x=\frac{1+y}{1+2z^2}$$
$$\begin{equation} x=y+3 \label{eq:xdef}
\end{equation}$$
In equation (\ref{eq:xdef}) we saw $\dots$ -->

```{r chunk2}
par(mar=c(2.5, 2.5, 0.5, 0.5)); plot(5:1)
```




```{r chunk1,fig.cap=fig$cap("five_to_one", "This is a plot of the numbers five to one")}
#remove all variables af data from global enviroment
#rm(list=ls())
suppressPackageStartupMessages(library(RCurl))
# load csv file from URL
x = getURL("https://raw.githubusercontent.com/juliuspeturg/Assignment1/master/fuel.csv")
fuel = read.csv(text=x,stringsAsFactors = FALSE)
summary(fuel)
attach(fuel)
par(mfrow=c(1,1))
plot(fpi~rtime,type = "l",main="Fuel price index in the US")
grid()
```



#### Question 2
For statinary time series the ACF will drop to zero very quickly however for nonstationary time series the ACF will taper off slowly. As we can see by looking at the ACF of the fuel price index the ACF tapers off slowly, from that we can conclude that the timeseries is nonstationary and therefor the mean of the series is not going to informe us a lot on the timeseries, see figure below for ACF of timeseries  
```{r eval=TRUE, echo=TRUE,message=FALSE,fig.cap=c("***Figure 1:***")}
#Generate timeseries object for fuelindex
fuel.ts = ts(data = fuel$fpi, start = c(1979,1),frequency = 12)
#evaluate the mean of the timeseries
cat(c("mean_fpi:",mean(fpi)))
acf(fuel.ts,main="ACF for fuel price index")
grid()
```

#### Question 3

```{r eval=TRUE, echo=TRUE,message=FALSE}
# Linear Model
mod <- lm(fpi~rtime)
summary(mod)
```

#### Question 4

```{r eval=TRUE, echo=TRUE,message=FALSE,fig.show='hold'}
# Trainingset
myts <- ts(fpi, start = c(1979,1), end =c(2003,12),frequency = 12)
# Testset
myts2 <- ts(fpi, start = c(2004,1), end =c(2004,12),frequency = 12)
require(forecast)
monthplot(myts)
seasonplot(myts)
par(mfrow=c(1,1))
fit1 <- stl(myts,s.window = "period")
#3 eða 4
plot(fit1)
```

#### Question 5

```{r eval=TRUE, echo=TRUE,message=FALSE}
fuel.ts <- ts(fuel$fpi, frequency=12, start=c(1979,1))

fuel_train <- window(fuel.ts, start=c(1979,1), end=c(2003,12))
fuel_test <- window(fuel.ts, start=c(2004,1), end=c(2004,12))


```

#### Question 6

```{r eval=TRUE, echo=TRUE,message=FALSE}
fit2 <- ses(fuel_train,h = 12, alpha = 0.1,exponential=TRUE)
f.fit2 <- forecast(fit2,12)
plot(f.fit2,ylab="fpi",xlab="time")
lines(fuel_train)
lines(fuel_test,col=2)
grid()
```

#### Question 7

```{r eval=TRUE, echo=TRUE,message=FALSE}
fit3 <- ses(fuel_train,h = 12, exponential=FALSE)
f.fit3 <- forecast(fit3,12)
plot(f.fit3,ylab="fpi",xlab="time")
lines(fuel_train)
lines(fuel_test,col=2)
grid()
```

#### Question 8

```{r eval=TRUE, echo=TRUE,message=FALSE}
fit4 <- HoltWinters(myts, beta = FALSE,gamma = FALSE)
f.fit4 <- forecast(fit4,12)
plot(f.fit4,ylab="fpi",xlab="time")
lines(fuel_train)
lines(fuel_test,col=2)
grid()
```

#### Question 9

```{r eval=TRUE, echo=TRUE,message=FALSE}
fit5 <- rwf(myts, h=12, drift=FALSE, fan=FALSE, lambda=NULL, biasadj=FALSE)
f.fit5 <- forecast(fit5,12)
plot(f.fit5,ylab="fpi",xlab="time")
lines(fuel_train)
lines(fuel_test,col=2)
grid()
```

------
### Assignment 2
#### Question 1

```{r eval=TRUE, echo=TRUE,message=FALSE}
library(RCurl)
x = getURL("https://raw.githubusercontent.com/juliuspeturg/Assignment1/master/simdata.csv")
simdata = read.csv(text=x,stringsAsFactors = FALSE,header=TRUE, sep="")

# A2-Q1: Plot the data

attach(simdata)

# á sama grafi
plot(y1, type="l",col='red', main="Simdata time series",
     xlab="Time", ylab="Data")
lines(y2,col='blue')
lines(y3,col='green')
legend('topright', legend=c("y1", "y2","y3"),
       col=c('red','blue','green'), lty=1, bty='y', cex=.85)
```

#### Question 2

```{r eval=TRUE, echo=TRUE,message=FALSE}
par(mfrow=c(3,2)) 
# par(mfrow=c(1,1)) 
acf(y1)
pacf(y1)
acf(y2)
pacf(y2)
acf(y3)
pacf(y3)
```

#### Question 3

```{r eval=TRUE, echo=TRUE,message=FALSE}
# A2-Q3: Identify the models using arima  <<< ATH BETUR

arima(y1)
arima(y2)
arima(y3)

auto.arima(y1) # líka hægt að nota auto.arima
```

#### Question 4

```{r eval=TRUE, echo=TRUE,message=FALSE}
ts.sim <- arima.sim(list(order = c(1,0,0), ar = 0.9), n = 250)
```

#### Question 5

```{r eval=TRUE, echo=TRUE,message=FALSE}
layout(matrix(c(1,1,2,3), 2, 2, byrow = TRUE))  
plot(ts.sim, main="Simulated time series ts.sim")
acf(ts.sim, main="ACF for ts.sim")
pacf(ts.sim, main="PACF for ts.sim")
```

#### Question 6

```{r eval=TRUE, echo=TRUE,message=FALSE}
acf.plot <- function(y){
  layout(matrix(c(1,1,2,3), 2, 2, byrow = TRUE)) 
  plot(y, main=c("Simulated time series",deparse(substitute(y))))
  acf(y, main=c("ACF for",deparse(substitute(y))))
  pacf(y,main=c("PACF for",deparse(substitute(y))))
}
```

#### Question 7

```{r eval=TRUE, echo=TRUE,message=FALSE,fig.show='hold'}
ts.sim1 <- arima.sim(n=250, list(ar=c(0.5)))        
ts.sim2 <- arima.sim(n=250, list(ar=c(-0.5)))
ts.sim3 <- arima.sim(n=250, list(ar=c(0.5,0.2)))
ts.sim4 <- arima.sim(n=250, list(ar=c(0.7)))
ts.sim5 <- arima.sim(n=250, list(ar=c(0.5,-0.2)))
ts.sim6 <- arima.sim(n=250, list(ma=c(0.5,0.5)))
ts.sim7 <- arima.sim(n=250, list(ma=c(-0.9)))
ts.sim8 <- arima.sim(n=250, list(ar=c(0.5), ma=c(-0.8)))

# Plotting the simulated series with the acf function
acf.plot(ts.sim1)
acf.plot(ts.sim2)
acf.plot(ts.sim3)
acf.plot(ts.sim4)
acf.plot(ts.sim5)
acf.plot(ts.sim6)
acf.plot(ts.sim7)
acf.plot(ts.sim8)
```

#### Question 8

```{r eval=TRUE, echo=TRUE,message=FALSE}
ts.sim9 <- arima.sim(list(order = c(1,1,1), ar = 0.7, ma=0.3), n = 250)

ts.sim9_diff1 <- diff(ts.sim9, differences=1) #first difference
plot.ts(ts.sim9_diff1, col=3)
lines(ts.sim9, col=2)
```

### **Assignment 3**
#### Question 1




### Assignment 4
#### Question 1

```{r eval=TRUE, echo=TRUE,message=FALSE}
rm(list=ls())
require(googleVis)
op <- options(gvis.plot.tag='chart')

require(RCurl)
x = getURL("https://raw.githubusercontent.com/juliuspeturg/Assignment1/master/gamma-visitolur.csv")
gamma <- read.csv(text=x, stringsAsFactors = FALSE, header=TRUE)

gamma$Date <- as.Date(gamma$Date, "%Y-%m-%d")
gamma <- na.omit(gamma)
names(gamma)[names(gamma)=="GAMMA.v.ed.sit.f6.lur"] <- "GAMMA Equity Index"

```

```{r,results='asis', tidy=FALSE,, fig.width = 7, fig.height = 7}
  A <- gvisAnnotationChart(gamma, datevar = 'Date', numvar = 'GAMMA Equity Index', date.format = "YY-mm-dd")
plot(A)  
```

```{r eval=TRUE,echo=FALSE}
## Set options back to original options
options(op)
```

#### Question 2

```{r eval=TRUE, echo=TRUE,message=FALSE}
max_row_training = max(which(gamma$Date=="2013-12-31"))
gamma_train <- head(gamma, n=max_row_training)
gamma_test <- tail(gamma, n=(nrow(gamma)-max_row_training))
gamma.ts <- ts(gamma$`GAMMA Equity Index`)
```

#### Question 3

```{r eval=TRUE, echo=TRUE,message=FALSE}
gamma.ts <- ts(gamma$`GAMMA Equity Index`)
gamma_train.ts <- ts(gamma_train$`GAMMA Equity Index`)
gamma_test.ts <- ts(gamma_test$`GAMMA Equity Index`) 

# fit_gamma <- ses(gamma_training$`GAMMA Equity Index`)
require(forecast)
fit_gamma <- HoltWinters(gamma_train.ts, beta = FALSE,gamma = FALSE)
f.fit_gamma <- forecast(fit_gamma, 20)
plot(f.fit_gamma)
accuracy(f.fit_gamma)
```

