---
title: "Project 2"
author: 
- "Júlíus Pétur Guðjohnsen"
- "Sigurður Jónsson"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output: 
  html_document: 
    fig_height: 4
    fig_width: 7
    fig_caption: yes
    self_contained: yes
    smart: no
---

```{r auto_num_functions, echo=FALSE}
# A function for generating captions and cross-references

fig <- local({
    i <- 0
    list(
        cap=function(refName, text, center=FALSE, col="black", inline=FALSE) {
            i <<- i + 1
            ref[[refName]] <<- i
            css_ctr <- ""
            if (center) css_ctr <- "text-align:center; display:inline-block; width:100%;"
            cap_txt <- paste0("<span style=\"color:", col, "; ", css_ctr, "\">Figure ", i, ": ", text , "</span>")
            anchor <- paste0("<a name=\"", refName, "\"></a>")
            if (inline) {
                paste0(anchor, cap_txt)    
            } else {
                list(anchor=anchor, cap_txt=cap_txt)
            }
        },
        
        ref=function(refName, link=FALSE, checkRef=TRUE) {
            
            ## This function puts in a cross reference to a caption. You refer to the
            ## caption with the refName that was passed to fig$cap() (not the code chunk name).
            ## The cross reference can be hyperlinked.
            
            if (checkRef && !refName %in% names(ref)) stop(paste0("fig$ref() error: ", refName, " not found"))
            if (link) {
                paste0("<A HREF=\"#", refName, "\">Figure ", ref[[refName]], "</A>")
            } else {
                paste0("Figure ", ref[[refName]])
            }
        },
        
        ref_all=function(){
            ## For debugging
            ref
        })
})

```

```{r setup, echo=FALSE}

## This chunk replaces the default hook for processing plots. It achieves the purposes,
## of laying out auto-numbered captions, but other functionality may be gone.

library(knitr)
knit_hooks$set(plot = function(x, options) {
    sty <- ""
    if (options$fig.align == 'default') {
        sty <- ""
    } else {
        sty <- paste0(" style=\"text-align:", options$fig.align, ";\"")
    }
    
    if (is.list(options$fig.cap)) {
        ## options$fig.cap is a list returned by the function fig$cap()
        str_caption <- options$fig.cap$cap_txt
        str_anchr <- options$fig.cap$anchor
    } else {
        ## options$fig.cap is a character object (hard coded, no anchor)
        str_caption <- options$fig.cap
        str_anchr <- ""
    }
    
    paste('<figure', sty, '>', str_anchr, '<img src="',
        opts_knit$get('base.url'), paste(x, collapse = '.'),
        '"><figcaption>', str_caption, '</figcaption></figure>',
        sep = '')
    
})
```


```{r read_fig_cap, echo=FALSE}
## This chunk will read through *this* Rmd file, and attempt to extract all of the 
## labels (not caption text) used for Figure captions. These labels are used
## as anchors, so scanning through the document now will allow us to create cross references
## before the caption actually appears. 

## Get the name of this Rmd file
rmdFn <- knitr::current_input()  # filename of input document

## Read lines and close connection
rmdCon <- file(rmdFn, open = "r")
rmdLines <- readLines(rmdCon)
close(rmdCon)

## Pull out all occurences of at least one back tick, followed 
## by any number of characters, followed by fig$cap (all on one line)
figscap_idx <- grep("`+(.*)fig\\$cap", rmdLines)
rmdLines <- rmdLines[figscap_idx]

## Get rid of everything up until the start of the caption label
## This presumes the caption label is the first argument of fig$cap()
## E.g., fig.cap = fig$cap("my_label", ...)
rmdLinesSansPre <- sub("(.*)fig\\$cap(.*?)[\"']", "", rmdLines)

## Identify everything up until the first quote
match_data <- regexpr("(.*?)[\"']", rmdLinesSansPre)

## Reduce the length by one, because we're not interested in the final quote
attr(match_data, "match.length") <- attr(match_data, "match.length") - 1

## Extract
fig_labels <- regmatches(rmdLinesSansPre, match_data, invert=FALSE)

if (length(fig_labels) > 0) {

    ## Test for duplicates
    if (anyDuplicated(fig_labels) > 0) stop("Duplicate caption labels detected")
    
    ## Create a named list of Figure numbers
    ref <- as.list(1:length(fig_labels))
    names(ref) <- fig_labels
}    

```




### **TASK 1**


The following code sample shows how to import the data needed for the assignement hosted on GitHub: https://raw.githubusercontent.com/juliuspeturg/Assignment2/.
For Task 1 we wrote function with three inputs: y,d,l where y is the input data, d is the number of differenceiation and l is the number of lags to plot. The output of the function is 4 plots: two ACF´s, PACF and a spectrum plot of the data.

```{r chunk1,eval=T}
suppressPackageStartupMessages(library(RCurl))
suppressPackageStartupMessages(library(ggplot2))
x = getURL("https://raw.githubusercontent.com/juliuspeturg/Assignment2/master/VEKS.csv")
VEKS = read.csv(text=x,stringsAsFactors = FALSE,header=TRUE)

attach(VEKS)
source("Multiplot.R")

xx <- qplot((1:length(HC.c)),HC.c,geom="line")+
    xlab("Time")+
    ylab("Heat Consumption")+
    ggtitle("Heat Consumption")
yy <- qplot((1:length(Ta.c)),Ta.c,geom="line")+
    xlab("Time")+
    ylab("Air Temperature")+
    ggtitle("Air Temperature")
multiplot(xx,yy,cols = 1)

```
`r fig$cap("heatVStemp", "Heat consumption and air temperature during the same one year period.", inline=TRUE)`

We notice that there are fluctuations in the heat consumption data on a small scale (day by day and week by week) and also a seasonal fluctuation, indicating more heat consumption during the winter and less during the summer when the ambient air temperature is higher.

```{r chunk1-1,eval=T}
acf.plot <- function(y,d,l){
  source("Multiplot.R")
  suppressPackageStartupMessages(require(ggplot2))
  suppressPackageStartupMessages(require(ggfortify))
  if(d!=0){
    y <- diff(y,difference=d)
  }
  else{}
  #----------Autocovariance--------#
  acf <- acf(y,type="covariance", na.action = na.pass ,plot = FALSE,lag.max=l)
  acfdf <- with(acf, data.frame(lag, acf))
  
  p1 <- ggplot(data = acfdf, mapping = aes(x = lag, y = acf)) +
    geom_hline(aes(yintercept = 0)) +
    geom_segment(mapping = aes(xend = lag, yend = 0))+
    xlab("lag")+
    ylab("ACF")+
    ggtitle("AutoCovariance")
  #----------AutoCorrelation--------#
  c_acf <- acf(y,type="correlation", na.action = na.pass ,plot = FALSE,lag.max=l)
  c_acfdf <- with(c_acf, data.frame(lag, acf))
  ci <- qnorm((1 + 0.95)/2)/sqrt(c_acf$n.used)
  p2 <- ggplot(data = c_acfdf, mapping = aes(x = lag, y = acf)) +
    geom_hline(aes(yintercept = 0)) +
    geom_segment(mapping = aes(xend = lag, yend = 0))+
    ggplot2::geom_hline(yintercept=c(-ci, ci), colour="blue", linetype="dashed")+
    xlab("lag")+
    ylab("ACF")+
    ggtitle("AutoCorrelation")
  
  #----------PACF--------#
  bpacf <- pacf(y,na.action = na.pass,plot = FALSE,lag.max=l)
  bpacfdf <- with(bpacf, data.frame(lag, acf))
  
  p3 <- ggplot(data = bpacfdf, mapping = aes(x = lag, y = acf)) +
    geom_hline(aes(yintercept = 0)) +
    geom_segment(mapping = aes(xend = lag, yend = 0))+
    ggplot2::geom_hline(yintercept=c(-ci, ci), colour="blue", linetype="dashed")+
    xlab("lag")+
    ylab("PACF")+
    ggtitle("Patial AutoCorrelation")
  #----------Spectrum--------#
  k = kernel("daniell", c(20,20))
  specvalues = spec.pgram(na.omit(y), k, taper=0, log = "no",plot = F)
  peak=specvalues$freq[which.max(specvalues$spec)]
  p4 <- autoplot(spec.pgram(na.omit(y), k, taper=0, log = "no",plot = FALSE))+
    geom_vline(xintercept = peak,colour = "red", linetype = "dotted")+
    ggtitle("Spectrum Plot")
  
  multiplot(p1,p2,p3,p4,cols = 2)
}

```

Now we can use our function to analyze the heat consumption time series.

```{r chunk2}
acf.plot(VEKS$HC.c,0,50)
```
`r fig$cap("plot1", "Plot of the two ACF´s (covariance and correlation) as well as the PACF and spectrum plot (difference=0).", inline=TRUE)`


As we can see in `r fig$ref("plot1",link=F)` the ACF tapers off very slowly. This indicates that the timeseries is non-stationary and we should differentiate to be able to extract any useful information from the plots. Next we estimate the series at difference=1 and take anther look at the outputs.


```{r chunk3}
acf.plot(VEKS$HC.c,1,50)
```
`r fig$cap("plot2", "Plot of the two ACF´s (covariance and correlation) as well as the PACF and spectrum plot (difference=1).", inline=TRUE)`


In `r fig$ref("plot2",link=F)` we can see a pattern emerge. We see some interesting spikes in lag 12, lag 24, lag 36 etc. Both the ACF and PACF show semi cyclical behavior but the largest spike is in lag 24. If we look at the spectrum plot of the series we see two distinct spikes with frequency of around 0.04 and 0.08 - corresponding to a period of 24 hours and 12 hours respectively. This tells us that the heat consumption goes up every 12 and 24 hours. Our best guess is that the heat consumption goes up in the morning and then again in the evening when people are spending time in their apartments/houses but goes down in between (people go to work or are sleeping).

### **TASK 2**

Intuitively we know that the data should have seasonality within every 24 hours, every week and also within every year.
When predicting for such a short time period (1 hour and 6 hours), we conclude that the 24 hour cycle is the most important seasonality cycle to look at, and therefore we change the data into a time series with frequency 24:

```{r Task2_1}

suppressPackageStartupMessages(require(forecast))
heat_con <- ts(na.omit(HC.c), frequency = 24)  

```
Let us examine the forecast for 6 hours ahead and first try a linear regression model using tslm():

```{r Task2_2}

fit_lm <- tslm(heat_con ~ trend + season)
accuracy(fit_lm)
fcast_lm <- forecast(fit_lm, h=6)
#summary(fit_lm)
autoplot(fcast_lm, include = 168)

```
`r fig$cap("LMforecast", "6 hour ahead forecast for the linear regression model.", inline=TRUE)`


After taking the first difference of the heat consumption time series, we inspect the ACF and PACF plots
and estimate that the heat consumption can be described with a ARIMA(3,1,1) process.
Using arima to predict with ARIMA(3,1,1):

```{r Task2_3}

fit1.arima <- arima(na.omit(HC.c), order=c(3,1,1))
#summary(fit1.arima)
accuracy(fit1.arima)
fcast1 <- forecast(fit1.arima,6)
autoplot(fcast1, include = 168)

```
`r fig$cap("ARIMAforecast", "6 hour ahead forecast using ARIMA(3,1,1).", inline=TRUE)`

Now we use the auto.arima() function from the forecast package in R. That gives a different model
than we estimated earlier, or ARIMA(4,1,4).

```{r Task2_4}

# auto.arima(na.omit(HC.c))  
# Gives ARIMA(4,1,4)
# Let's now use this input to predict:
fit2.arima <- arima(na.omit(HC.c), order=c(4,1,4))
#summary(fit2.arima)
accuracy(fit2.arima)
fcast2 <- forecast(fit2.arima,6)
autoplot(fcast2, include = 168)

```
`r fig$cap("autoarima(4,1,4)forecast", "6 hour ahead forecast using ARIMA(4,1,4).", inline=TRUE)`

Lets examine the forecast for 1 hour ahead with the same methods as above:



### **TASK 3**

For this task we did simmilar things as in task 1: made a function to plot the cross correlation of the possible measurements that might be used to aid in forecasting of the heat consumption. The measurements for air temperature, wind speed and global radiation were used. The cross correlation of the three measurements and the heat consumption were evaluated as can be seen below.

```{r chunk4,eval=T}
CCF.plot <- function(d,l){
  source("Multiplot.R")
  suppressPackageStartupMessages(require(forecast))  
  suppressPackageStartupMessages(require(ggplot2))
  suppressPackageStartupMessages(require(ggfortify))
  
  HeatConsumption <- VEKS$HC.c
  Air_Temperature <- VEKS$Ta.c
  WindSpeed       <- VEKS$W.c 
  GlobalRadiation <- VEKS$GR.c
  
  if(d!=0){
    HeatConsumption <- diff(VEKS$HC.c,difference=d)
    Air_Temperature     <- diff(VEKS$Ta.c,difference=d)
    WindSpeed       <- diff(VEKS$W.c, difference=d) 
    GlobalRadiation <- diff(VEKS$GR.c,difference=d)
  }
  else{}
  
  ccf1 <- autoplot(stats::ccf(HeatConsumption,Air_Temperature,plot=FALSE,na.action=na.pass,lag.max = l))+
    geom_vline(xintercept = 0,colour = "red", linetype = "dotted")+
    ggtitle("Air temp. & Heat Consumption")


  ccf2 <- autoplot(stats::ccf(HeatConsumption,GlobalRadiation,plot=FALSE,na.action=na.pass,lag.max = l))+
    geom_vline(xintercept = 0,colour = "red", linetype = "dotted")+
    ggtitle("Global rad. & Heat Consumption")

  
  ccf3 <- autoplot(stats::ccf(HeatConsumption,WindSpeed,plot=FALSE,na.action=na.pass,lag.max = l))+
    geom_vline(xintercept = 0,colour = "red", linetype = "dotted")+
    ggtitle("Wind & Heat Consumption")

  
  multiplot(ccf1,ccf2,ccf3,cols=1)
}

```

The input to the plotfunction is d and l for the lag.max, the difference to be used on the data. So lets see what the cross correlation looks like at difference=0:

```{r chunk5}
CCF.plot(0,20)
```
`r fig$cap("plot3", "Plot of the cross correlation between the heat consumption and air temperature(top), windspeed (middle) and global radiation(bottom) with difference=0. The red dotted line is Lag=0.", inline=TRUE)`

As we can see in `r fig$ref("plot3")` there is not much information to be extracted from these plots. Again that is due to the series being non-stationary so we plot again, now with difference=1.

```{r chunk6}
CCF.plot(1,20)
```
`r fig$cap("plot4", "Plot of the cross correlation between the heat consumption and air temperature(top), windspeed (middle) and global radiation(bottom) with difference=1. The red dotted line is Lag=0.", inline=TRUE)`

From `f fig$ref("plot4")` we believe that there might be some indicator of a lagged effect in lag 0 and lag 1 for the air temperature and windspeed and some at lag 1 for global radiation. Now it should be interesting to see what pre-whitening will do to the data. Lets start by looking at air temperature:

```{r chunk7,eval=F,echo=F}
HeatConsumption <- VEKS$HC.c
Air_Temperature <- VEKS$Ta.c
WindSpeed       <- VEKS$W.c 
GlobalRadiation <- VEKS$GR.c

x1 <- ggAcf(Air_Temperature,na.action=na.pass)+ggtitle("Acf for Air Temp. diff 0")
x2 <- ggPacf(Air_Temperature,na.action=na.pass)+ggtitle("Pacf for Air Temp. diff 0")
x3 <- ggAcf(diff(Air_Temperature,difference=1),na.action=na.pass)+ggtitle("Acf for Air Temp. diff 1")
x4 <- ggPacf(diff(Air_Temperature,difference=1),na.action=na.pass)+ggtitle("Pacf for Air Temp. diff 1")

multiplot(x1,x3,x2,x4,cols=2)
```

```{r chunk8,eval=T,echo=F}
pw.air_temp<-auto.arima(VEKS$Ta.c)
#pw.temp = ARIMA(3,1,2) with drift
pw.Hc<-residuals(arima(VEKS$HC.c,order=c(3,1,2),fixed=c(pw.air_temp$coef[1:5])))
x <- pw.air_temp$residuals
y <- pw.Hc
ggCcf(x,y,lag.max=25)+ggtitle("Pre-whitening series: Air temp. & Heat consumption")+
    geom_vline(xintercept = 0,colour = "red", linetype = "dotted")+
    geom_vline(xintercept = c(-12,12),colour = "red", linetype = "dotted")+
    geom_vline(xintercept = c(-24,24),colour = "red", linetype = "dotted")
```
`r fig$cap("plot5", "Pre-whitening - here we can see a strong indicator in lag 1.", inline=TRUE)`


```{r chunk9,eval=T,echo=F}
pw.wind<-auto.arima(VEKS$W.c)
#pw.wind = ARIMA(4,1,2)
pw.Hc<-residuals(arima(VEKS$HC.c,order=c(4,1,2),fixed=c(pw.wind$coef)))
x <- pw.wind$residuals
y <- pw.Hc
ggCcf(x,y,lag.max=25)+ggtitle("Pre-whitening - Wind & Heat consumption")+
    geom_vline(xintercept = 0,colour = "red", linetype = "dotted")+
    geom_vline(xintercept = c(-12,12),colour = "red", linetype = "dotted")+
    geom_vline(xintercept = c(-24,24),colour = "red", linetype = "dotted")
```
`r fig$cap("plot6", "Pre-whitening - strong indicator in lag 1", inline=TRUE)`


```{r chunk10,eval=T,echo=F}
pw.radiation<-auto.arima(VEKS$GR.c)
#pw.radiation = ARIMA(3,1,2)
pw.Hc<-residuals(arima(VEKS$HC.c,order=c(3,1,2),fixed=c(pw.radiation$coef)))
x <- pw.radiation$residuals
y <- pw.Hc
ggCcf(x,y,lag.max=25)+ggtitle("Pre-whitening series: Radiation & Heat consumption")+
    geom_vline(xintercept = 0,colour = "red", linetype = "dotted")+
    geom_vline(xintercept = c(-12,12),colour = "red", linetype = "dotted")+
    geom_vline(xintercept = c(-24,24),colour = "red", linetype = "dotted")
```
`r fig$cap("plot7", "Pre-whitening. here we can strong indicator in lag 0.", inline=TRUE)`
